# Data sources

II. Data sources
Describe the data sources: who is responsible for collecting the data? How is it collected? If there were a choice of options, explain how you chose.

Provide some basic information about the dataset: types of variables, number of records, etc.

Describe any issues / problems with the data, either known or that you discover.

(suggested: approximately 1 page)

III. Data transformation
Describe the process of getting the data into a form in which you could work with it in R. If your code does not lend itself to being including in the .Rmd  file, provide a link to the folder or file(s) that contain(s) that code.

(suggested: approximately 1/2 page)

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
data1.0 <- read.csv("1 visitor trend-不分state.csv")[-c(7:11),]
data2.0 <- read.csv("2 visitor trend-分state.csv")[-c(301:305),]
data2.10<- read.csv("2.1 visitor trend- state plus dest.csv")[-c(1051:1054),]
data3.0 <- read.csv("3 destination.csv")[-c(29:32),]
data4.0 <- read.csv("4 Q2主要数据.csv")[-c(217:221),]
```