
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
# Data transformation

"Describe the process of getting the data into a form in which you could work with it in R. If your code does not lend itself to being including in the .Rmd  file, provide a link to the folder or file(s) that contain(s) that code."

```{r}
# possible better format
data1.0 = read.csv("data/raw/1 visitor trend-notbystate.csv")
data2.0 = read.csv("data/raw/2 visitor trend-bystate.csv")
data3.0 = read.csv("data/raw/3 destination.csv")
data4.0 = read.csv("data/raw/4 Q2 main data.csv")

#跑完所有预处理最后加上
write.csv(d1,"data/clean/d1.csv", row.names = FALSE)
#....
#下次直接read.csv("data/clean/d1.csv")
```

```{r}
# from chapter 2
data1.0 <- read.csv("1 visitor trend-不分state.csv")[-c(7:11),]
data2.0 <- read.csv("2 visitor trend-分state.csv")[-c(301:305),]
data2.10<- read.csv("2.1 visitor trend- state plus dest.csv")[-c(1051:1054),]
data3.0 <- read.csv("3 destination.csv")[-c(29:32),]
data4.0 <- read.csv("4 Q2主要数据.csv")[-c(217:221),]
```

To tidy our data, firstly we chose to convert all the blank observations into NAs using a helper function "empty_as_na". Then, we deleted redundant characters in the names of columns, for instance, all the "X"s in front of the years like "X2019". As there are a lot of missing values through the years, we may especially concern about the data since 2019 when analyzing the first two problems. We kept the data in year 2019 using "filter" command in tidyverse package. Our code is shown below

????????????????? how "provide a link to the folder or file(s) that contain(s) that code"

```{r}
library(tidyverse)
library(dplyr)

empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    #ifelse(as.character(x)!="", x, NA)
    ifelse(x!="", x, NA)
}


data1<- data1.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2<- data2.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2.1<- data2.10 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data3<- data3.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data4<- data4.0 %>%
  pivot_longer(cols = -c("Group", "Indicator", "Units"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

d1<- tibble(data1[,-c(5:21)][,-c(3)][,c(1,2,3,8)])
d2<- tibble(data2[,-c(5:21)]) 
d2.1<- tibble(data2.1)[,-c(5:21)][,c(1,2,3,4,8)]
d4<- tibble(data4.0[,-c(4:19)][,c(1, 2, 8)])
```




