


# Data transformation


```{r}
# from chapter 2
data1.0 <- read.csv("data/raw/1 visitor trend-不分state.csv")[-c(7:11),]
data2.0 <- read.csv("data/raw/2 visitor trend-分state.csv")[-c(301:305),]
data2.10<- read.csv("data/raw/2.1 visitor trend- state plus dest.csv")[-c(1051:1054),]
data3.0 <- read.csv("data/raw/3 destination.csv")[-c(29:32),]
data4.0 <- read.csv("data/raw/4 Q2主要数据.csv")[-c(217:221),]
```

To tidy our data, firstly we chose to convert all the blank observations into NAs using a helper function "empty_as_na". Then, we deleted redundant characters in the names of columns, for instance, all the "X"s in front of the years like "X2019". As there are a lot of missing values through the years, we may especially concern about the data since 2019 when analyzing the first two problems. We kept the data in year 2019 using "filter" command in tidyverse package.

Except for changing column names, we also tried various ways to tidy our data. We mainly used pivot_longer, pivot_wider, mutate and filter to transform our data into the form that could be our best using. Eventually, the data is transformed pretty neat and is stored in the "clean" folder in our github respiratory "HawaiiTravelingData".

The code of data transformation could be found in the link: https://github.com/Weijiaaaa/5293_r/blob/main/Final%20project%20Ch3%20data%20transformation

```{r}
library(tidyverse)
library(dplyr)

empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    #ifelse(as.character(x)!="", x, NA)
    ifelse(x!="", x, NA)
}


data1<- data1.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2<- data2.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2.1<- data2.10 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data3<- data3.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data4<- data4.0 %>%
  pivot_longer(cols = -c("Group", "Indicator", "Units"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

d1<- tibble(data1[,-c(5:21)][,-c(3)][,c(1,2,3,8)])
d2<- tibble(data2[,-c(5:21)]) 
d2.1<- tibble(data2.1)[,-c(5:21)][,c(1,2,3,4,8)]
d4<- tibble(data4.0[,-c(4:19)][,c(1, 2, 8)])
```


```{r}

#跑完所有预处理最后加上
write.csv(d1,"data/clean/d1.csv", row.names = FALSE)
write.csv(d2,"data/clean/d2.csv", row.names = FALSE)
write.csv(data3,"data/clean/d3.csv", row.names = FALSE)
write.csv(d4,"data/clean/d4.csv", row.names = FALSE)
write.csv(d2.1,"data/clean/d2.1.csv", row.names = FALSE)

#....
#下次直接read.csv("data/clean/d1.csv")
```


