
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Results


V. Results
You have a lot of freedom to choose what to do, as long as you restrict yourselves to exploratory techniques (rather than modeling / prediction approaches). In addition, your analysis must be clearly documented and reproducible.

Provide a short nontechnical summary of the most revealing findings of your analysis written for a nontechnical audience. Take extra care to clean up your graphs, ensuring that best practices for presentation are followed, as described in the  audience ready style section  below.

Use subheadings as appropriate. See  Todd Schneider’s blog posts  for examples of thoughtful, informative subheadings.

The number of graphs will vary by project; we suggest a target of 10. (A plot with multiple facets counts as 1 graph).  If you go over, make sure they are high quality and include commentary.  A project with 10 good graphs plus commentary will receive a better grade than a project with 10 good graphs and 10 mediocre ones.


```{r}
# from chapter 2&3
data1.0 <- read.csv("1 visitor trend-不分state.csv")[-c(7:11),]
data2.0 <- read.csv("2 visitor trend-分state.csv")[-c(301:305),]
data2.10<- read.csv("2.1 visitor trend- state plus dest.csv")[-c(1051:1054),]
data3.0 <- read.csv("3 destination.csv")[-c(29:32),]
data4.0 <- read.csv("4 Q2主要数据.csv")[-c(217:221),]

library(tidyverse)
library(dplyr)

empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    #ifelse(as.character(x)!="", x, NA)
    ifelse(x!="", x, NA)
}


data1<- data1.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2<- data2.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

data2.1<- data2.10 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data3<- data3.0 %>% 
  mutate_each(funs(empty_as_na)) %>%
  pivot_longer(cols = -c("Market", "Indicator", "Units", "Destination"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")


data4<- data4.0 %>%
  pivot_longer(cols = -c("Group", "Indicator", "Units"),
               names_to = "Year",
               values_to = "number")  %>%
  mutate(Year = substr(Year, 2, 1000)) %>%
  pivot_wider(names_from = "Year",
              values_from = "number")

d1<- tibble(data1[,-c(5:21)][,-c(3)][,c(1,2,3,8)])
d2<- tibble(data2[,-c(5:21)]) 
d2.1<- tibble(data2.1)[,-c(5:21)][,c(1,2,3,4,8)]
d4<- tibble(data4.0[,-c(4:19)][,c(1, 2, 8)])
```

```{r}
d4.1<- d4 %>%
  filter(Group != "Visitor from US") %>%
  na.omit() %>%
  pivot_wider(names_from = Indicator, values_from = X2019) 
  
d4.1 %>%
  mutate(Group = factor(Group, levels = c("Oahu visitors", 
                                          "Maui visitors",
                                          "Hawaii ISL visitors",
                                          "Kauai visitors",
                                          "Lanai visitors",
                                          "Molokai visitors"))) %>%
  ggplot(aes(x = Group, y = `Avg. age of party head`)) +
  geom_col(fill = "cornflowerblue") +
  coord_flip() 


d4.1[,c(1,3,4)] %>%
  pivot_longer(c(`First timer (%)`, `Repeaters (%)`), 
               names_to ="Times of come", 
               values_to = "percentage") %>%
  ggplot(aes(x = Group, y = percentage, fill = `Times of come`)) +
  geom_col() 



purpose<- cbind(d4.1[,c(1)], d4.1[c(grep("Purpose",colnames(d4.1)))])%>%
  pivot_longer(c(2:16), 
               names_to ="Purpose", 
               values_to = "percentage") %>%
  mutate(Purpose = substr(Purpose, 10, 1000)) %>%
  mutate(Purpose = gsub('.{4}$', '', Purpose)) %>%
  filter(Purpose %in% c("Vacation", "Honeymoon/Get married", "Visit Friends/Relative")) %>%
  pivot_wider(names_from = Purpose, 
              values_from = percentage) %>%
  mutate(OtherPurpose = 100 - Vacation - `Honeymoon/Get married` - `Visit Friends/Relative`) %>%
  pivot_longer(c(2:5),
               names_to = "Purpose",
               values_to = "percentage")

purpose %>%
  ggplot(aes(x = Group, y = percentage, fill = Purpose)) +
  geom_col()  +
  theme(axis.text.x = element_text(angle = 10, vjust = 0.5, hjust=1))



Accom<- cbind(d4.1[,c(1)], d4.1[c(grep("Accom",colnames(d4.1)))]) %>%
  pivot_longer(c(2:14), 
               names_to ="Accomodation", 
               values_to = "percentage") %>%
  mutate(Accomodation = substr(Accomodation, 8, 1000)) %>%
  mutate(Accomodation = gsub('.{4}$', '', Accomodation)) %>%
  filter(Accomodation %in% c("Hotel only", "Condo", "Friends/Relatives", "Rental house")) %>%
  pivot_wider(names_from = Accomodation, 
              values_from = percentage) %>%
  mutate(OtherAccomodation = 100 - `Hotel only` - `Condo` - `Friends/Relatives` - `Rental house`) %>%
  pivot_longer(c(2:6),
               names_to = "Accomodation",
               values_to = "percentage")

Accom %>%
  ggplot(aes(x = Group, y = percentage, fill = Accomodation)) +
  geom_col() +
    theme(axis.text.x = element_text(angle = 10, vjust = 0.5, hjust=1))

library(treemap)

d3<- tibble(data3[,-c(5:29)][,c(1,2,3,4,9)])
colnames(d3)[5] <- 'Year_2019'
d3$Year_2019<- as.numeric(gsub(",","", d3$Year_2019))

d3 %>%
  filter(Indicator == "Visitor arrivals") %>%
  select( -c(Market, Units)) %>%
  treemap::treemap(index=c("Destination"),
                   vSize="Year_2019",
                   vColor="Year_2019",
                   type="value",
                   title="Number of visitors to different islands",  
                   format.legend = list(scientific = FALSE, big.mark = " ")) 
  
d3 %>%
  filter(Indicator == "Length of stay") %>%
  select( -c(Market, Units)) %>%
  treemap::treemap(index=c("Destination"),
                   vSize="Year_2019",
                   vColor="Year_2019",
                   type="value",
                   title="Length of stay for visitors to different islands",  
                   format.legend = list(scientific = FALSE, big.mark = " ")) 




```